### 2023-11-22 README.Rmd
update readme

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
# Please put your title here to include it in the file below.
Title <- "ROF Camera Trap Data Analysis - Preliminary Report"
```

# ROF Camera Trap Data Analysis Research Compedium

<!-- [![Binder](https://mybinder.org/badge_logo.svg)](https://mybinder.org/v2/gh/StewartWILDlab/rofcamtrap/main?urlpath=rstudio) -->

Prerequisites:

  - docker
  - nvidia-container-toolkit

## Main Workflow

1. At the terminal, clone this repository.

```{bash eval=FALSE}
git clone https://github.com/StewartWILDlab/rofcamtrap
```

2. Build the docker image of the computing environment, by running the build script. This will also build the apptainer image needed on HPC.

```{bash eval=FALSE}
rofcamtrap/dockerfiles/build.sh
```

3. Run docker image, with the proper volumes. We currently have two separate storage volumes, for each of the two camera retrievals that took place. We also make sure to hook the `rofcamtrap` folder as a volume.

```{bash eval=FALSE}
# -e DISABLE_AUTH=true --shm-size 50G
docker run \
  -v "$(pwd):/workspace/rofcamtrap" \
  -v "/media/vlucet/TrailCamST/TrailCamStorage:/workspace/storage/TrailCamStorage" \
  -v "/media/vlucet/TrailCamST/TrailCamStorage_2:/workspace/storage/TrailCamStorage_2" \
  --gpus all \
  -it rofcamtrap
```

4. Activate ENV, then run mega detector on images using the bash script.

```{bash eval=FALSE}
mamba activate cameratraps-detector
rofcamtrap/scripts/bash/camtrap.sh \
  -b "/workspace/git" \
  -s "/workspace/storage/TrailCamStorage_2" \
  -m "/workspace/models/md_v5a.0.0.pt" \
  -o "/workspace/rofcamtrap/1_MegaDetector/0_outputs/TrailCamStorage_2" \
  md
```

5. Run the repeat detector using the bash script, and remove all instances of trues positives.

```{bash eval=FALSE}
rofcamtrap/scripts/bash/camtrap.sh \
  -b "/workspace/git" \
  -s "/workspace/storage/TrailCamStorage_2" \
  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs/TrailCamStorage_2" \
  repeat-detect
```

6. Patch MD's output with the repeat detect results.

```{bash eval=FALSE}
rofcamtrap/scripts/bash/camtrap.sh \
  -b "/workspace/git" \
  -s "/workspace/storage/TrailCamStorage_2" \
  -i "/workspace/rofcamtrap/1_MegaDetector/0_outputs/TrailCamStorage_2" \
  -o "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/TrailCamStorage_2" \
  repeat-remove
```

7. Optionally, write out the visualizations of the detections.

```{bash eval=FALSE}
rofcamtrap/scripts/bash/camtrap.sh \
  -b "/workspace/git" \
  -s "/workspace/storage/TrailCamStorage_2" \
  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/TrailCamStorage_2" \
  -o "/workspace/rofcamtrap/1_MegaDetector/2_visualize/TrailCamStorage_2" \
  viz
```

8. Convert to coco and ls [and 3rd format?].

```{bash eval=FALSE}
mamba deactivate 
cd mdtools
poetry shell
cd ../
rofcamtrap/scripts/bash/camtrap.sh \
  -b "/workspace/git" \
  -s "/workspace/storage/TrailCamStorage_2" \
  -i "/workspace/rofcamtrap/1_MegaDetector/1_outputs_no_repeats/TrailCamStorage_2" \
  -o "/workspace/rofcamtrap/2_LabelStudio/0_inputs/TrailCamStorage_2" \
  repeat-convert
```

9. Crop annotations.

```{bash eval=FALSE}
rofcamtrap/scripts/bash/camtrap.sh \
  -b "/workspace/git" \
  -s "/workspace/storage/TrailCamStorage_2" \
  -i "/workspace/rofcamtrap/2_LabelStudio/0_inputs/TrailCamStorage_2" \
  crop
```

## Label studio outputs

1. Enter the container running the label studio app and output ths number of projects

```{bash eval=FALSE}
docker exec -it label-studio-app-1 bash
curl -X GET http://localhost:8080/api/projects/ -H 'Authorization: Token 3135fdd1f4a5b9b3630b69011ec4d70e7800c41d' -o files/outputs/project_counts.json
```

2. On the instance outside the container, run the Python script to extract projects id

```{python eval=FALSE, python.reticulate = FALSE}
import json

with open('data/outputs/project_counts.json', 'r') as file:
  data = json.load(file)

ids = [data['results'][i]['id'] for i in range(len(data['results']))]

with open('data/outputs/project_ids.txt', 'w') as file:
  for the_id in ids:
        file.write(str(the_id) + '\n')
```

3. Back in the container

```{bash eval=FALSE}
arr=(170 86 85 84 83 82 81 80 79 78 77 76 75 74 73 72 65 64 63 62 61 60 59 58 57 55 54 53 52 51 50 49 48 47 46 44 43 42 41 40 39 37 35 33 32 31 30 28 25 23 21 20 19 18 17 16 15 13 10); 
for id in ${arr[@]}; do         
  # url="http://localhost:8080/api/projects/${id}/export?exportType=JSON&download_all_tasks=true";
  url="http://localhost:8080/api/projects/${id}/export?exportType=JSON";
  curl -X GET "$url" -H 'Authorization: Token 3135fdd1f4a5b9b3630b69011ec4d70e7800c41d'\
    -o "files/outputs/ls/output_file_$id.json";
done
```

4. Check for file numbers

```{bash eval=FALSE}
ls data/outputs/ls # 59 nov 21
```

5. Copy to local machine

```{bash eval=FALSE}
scp -i ssh_key/arbutus_def_fstewart_prod 'ubuntu@206.12.94.17:~/data/outputs/ls/*' rofcamtrap/2_LabelStudio/1_outputs_downloaded/
```

6. Process those outputs

```{bash eval=FALSE}
rofcamtrap/scripts/bash/camtrap.sh \
  -i "/workspace/rofcamtrap/2_LabelStudio/1_outputs_downloaded/" \
  -o "/workspace/rofcamtrap/2_LabelStudio/2_outputs_processed" \
  post
```

## Classifier training workflow

On beluga, we use the apptainer image instead.

```{bash eval=FALSE}
. /workspace/conda/etc/profile.d/conda.sh 
. /workspace/conda/etc/profile.d/mamba.sh
PATH="$PATH:$HOME/.local/bin"
mamba activate cameratraps-detector
rofcamtrap/scripts/bash/camtrap.sh -b "/workspace/git" -s "/workspace/storage/my_passport_images" -m "/workspace/models/md_v5a.0.0.pt" md

$ mkdir -p /scratch/$USER/apptainer/{cache,tmp}
$ export APPTAINER_CACHEDIR="/scratch/$USER/apptainer/cache"
$ export APPTAINER_TMPDIR="/scratch/$USER/apptainer/tmp"

salloc --time=00:05:00 --mem=4G --ntasks=1 --gpus-per-task=1 --cpus-per-task=1 --account=rrg-fstewart

apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap"  -B "/media/vlucet/TrailCamST/TrailCamStorage:/workspace/storage/TrailCamStorage"  -B "/media/vlucet/My Passport/Images:/workspace/storage/my_passport_images" rofcamtrap.sif

apptainer shell --nv -C -B "$(pwd):/workspace/rofcamtrap" -B "/home/vlucet/projects/rrg-fstewart/vlucet:/workspace/project/" rofcamtrap.sif
```

### Species classifier

We need to download the models... TBA

### False detections classifier

## LabelStudio instance setup

## Labelme? WildTrax?

## Reports

The **reports** directory currently contains the `GnC_report` folder wich is structured as such:

  - [:file\_folder: paper](/analysis/paper): Quarto source document
    for manuscript. Includes code to reproduce the figures and tables
    generated by the analysis. It also has a rendered version,
    `paper.docx`, suitable for reading (the code is replaced by figures
    and tables in this file)
  - [:file\_folder: data](/analysis/data): Data used in the analysis.
  - [:file\_folder: figures](/analysis/figures): Plots and other
    illustrations
  - [:file\_folder: supplementary-materials](/analysis/supplementary-materials):
    Supplementary materials including notes and other documents
    prepared and collected during the analysis.

## How to run in your browser or download and run locally

This research compendium has been developed using the statistical programming
language R. To work with the compendium, you will need
installed on your computer the [R software](https://cloud.r-project.org/)
itself and optionally [RStudio Desktop](https://rstudio.com/products/rstudio/download/).

You can download the compendium as a zip from from this URL:
[master.zip](/archive/main.zip). After unzipping:
- open the `.Rproj` file in RStudio
- run `devtools::install()` to ensure you have the packages this analysis depends on (also listed in the
[DESCRIPTION](/DESCRIPTION) file).
- finally, open `analysis/paper/paper.Rmd` and knit to produce the `paper.docx`, or run `rmarkdown::render("analysis/paper/paper.qmd")` in the R console

### Licenses

TBD

<!-- **Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse -->

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

### How to cite

Please cite this compendium as:

> Lucet, Valentin; Stewart, Frances et al., (`r format(Sys.Date(), "%Y")`). _Compendium of R code and data for `r Title`_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### Notes

```{bash eval=FALSE}
for FILE in project*
    mdtools postprocess --write-csv $FILE
end
```

<!-- This repository contains the data and code for our paper:

> Authors, (YYYY). _`r Title`_. Name of journal/book <https://doi.org/xxx/xxx>

Our pre-print is online here:

> Authors, (YYYY). _`r Title`_. Name of journal/book, Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx> -->