### 2024-01-28 reports/GnC_report/paper/paper_test.qmd
add test version

---
title: "Using Camera Traps and Computer Vision to Quantify Wildlife Diversity and Co-Occurrence across \nOntario’s Far North"
subtitle: "Prepared for: Canadian Wildlife Service, Canadian Forest Service, and Wilfrid Laurier University"
author:
  - name: Valentin Lucet
    email: valentin.lucet@gmail.com
    orcid: 0000-0003-0268-818X
    affiliation: 
      - ref: wlu
affiliations:
  - id: wlu
    name: Wilfrid Laurier University
title-block-published: "Last updated"  
date: now
date-format: long
format:
  html: default
  # pdf:
  #   toc: true
  #   number-sections: true
  #   colorlinks: true
  #   extra_dependencies: ["float"]
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
bibliography: references.bib
abstract: >
  Test.
---

## Test