### 2024-04-29 analysis/documents/1-data-preparation.qmd
updates

---
title: "Data Processing"
author:
  - name: Valentin Lucet
    email: valentin.lucet@gmail.com
    orcid: 0000-0003-0268-818X
    affiliation: 
      - ref: wlu
affiliations:
  - id: wlu
    name: Wilfrid Laurier University
title-block-published: "Last updated"  
date: now
date-format: long
format:
  html: default
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
    extra_dependencies: ["float"]
format-links: false
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-dpi: 600
# bibliography: references.bib
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
backup_options <- options()
options(scipen = 1, digits = 2)
do_eval <- FALSE
```

In this script, we format the data needed for further analyses.

#### Process detections

```{r eval=do_eval}
#| label: process-detections

detections <- rofcamtrap::process_detections(
  input_dir = here::here("2_LabelStudio", "0_inputs"))

saveRDS(detections,  here::here("data", "objects" ,"detections.rds"))
```

#### Process annotations

```{r eval=do_eval}
#| label: process-annotationa

anns_wide <- rofcamtrap::process_annotations(
  input_dir = here::here("2_LabelStudio", "2_outputs_processed/"))

saveRDS(anns_wide, here::here("data", "objects", "annotations_wide_treated.rds"))

# Filter for non empty images
anns_wide_noempty <- anns_wide |>
  dplyr::filter(!is.na(to_name))

saveRDS(anns_wide_noempty, here::here("data", "objects", "annotations_wide_noempty.rds"))

# anns_wide_noempty <- readRDS(here::here("data", "objects", "annotations_wide_noempty.rds"))
readr::write_csv(anns_wide_noempty, here::here("data", "tabular", "annotations_wide_noempty.csv"))
```

#### Process exif data

```{r eval=do_eval}
#| label: process-exif

exif <- readr::read_csv(here::here("data", "tabular", "all_exif.csv")) |>
  janitor::clean_names() |>
  dplyr::mutate(image_id = stringr::str_replace_all(source_file, "/", "_")) |>
  dplyr::mutate(image_id = stringr::str_replace_all(image_id, ".JPG", "")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/media/vlucet/TrailCamST/TrailCamStorage/")) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage_2/")) |>
  dplyr::mutate(source_file = stringr::str_remove(source_file, "/workspace/storage/TrailCamStorage/")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_media_vlucet_TrailCamST_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_2_")) |>
  dplyr::mutate(image_id = stringr::str_remove(image_id, "_workspace_storage_TrailCamStorage_")) |> 
  dplyr::mutate(image_id = strsplit(image_id, "_") |> 
                  purrr::map(~paste0(.x[2:length(.x)], collapse = "_")) |> 
                  unlist()) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage/"))) |> 
  dplyr::mutate(source_file = stringr::str_remove(source_file, stringr::fixed("/media/vlucet/TrailCamST/TrailCamStorage_2/"))) #|> 
#   dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
#                                                tz = "Canada/Eastern"),
#                 year = lubridate::year(date_time),
#                 month = lubridate::month(date_time, label = TRUE),
#                 day = lubridate::day(date_time),
#                 ymd = paste(year, month, day, sep = "_"))
# 
# exif[
#   exif$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
# exif[
#   exif$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"

readr::write_csv(exif, here::here("data", "tabular", "exif_processed.csv"))
```

#### Load data

```{r eval=do_eval}
#| label: load-data
detections <- readRDS(here::here("data", "objects", "detections.rds"))
anns_wide <- readRDS(here::here("data", "objects", "annotations_wide_treated.rds"))
anns_wide_noempty <- readRDS(here::here("data", "objects", "annotations_wide_noempty.rds"))
exif <- readr::read_csv(here::here("data", "tabular", "exif_processed.csv"))
```

#### Process spatial information

```{r eval=do_eval}
#| label: process-cams-sf

RoF_CameraDeploymentLog_WLU <- 
  suppressWarnings(readxl::read_excel(here::here("data-raw", "tabular", "RoF_CameraDeploymentLog.xlsx"),
                                      guess_max = 10)) |>
  janitor::clean_names() |> 
  dplyr::mutate(date_retrieved = as.Date.character(date_retrieved))

cams_sf <- RoF_CameraDeploymentLog_WLU |> 
  dplyr::rename(longitude = x, latitude = y) |> 
  dplyr::filter(!is.na(latitude)) |> 
  dplyr::filter(!is.na(longitude)) |> 
  sf::st_as_sf(coords = c("longitude","latitude"), crs = "EPSG:4326", remove = F) |> 
  dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))

coords <- cams_sf |> 
  sf::st_coordinates() |> 
  as.data.frame()

cams_sf_coords <- dplyr::bind_cols(cams_sf, coords) |>
  dplyr::mutate(
    month_year_deployed = paste(lubridate::month(date_deployed, label = T, abbr = T), 
                                lubridate::year(date_deployed)), 
    month_year_retrieved = paste(lubridate::month(date_retrieved, label = T, abbr = T), 
                                 lubridate::year(date_retrieved)), 
    deployment_group = paste0(month_year_deployed, ifelse(
      month_year_retrieved == "NA NA", "", paste(" -", month_year_retrieved))))

saveRDS(cams_sf_coords, here::here("data", "objects", "cams_sf_coords.rds"))
```

```{r eval=do_eval}
#| label: process-zones
zones <- sf::st_read(here::here("data-raw","spatial", "ecozones", "EcoZone" ,"EcoZone.shp"),
              quiet = TRUE) |> 
  sf::st_transform(crs = sf::st_crs(cams_sf))
zones_filtered <- zones |>
  sf::st_filter(cams_sf_coords)

saveRDS(zones_filtered, here::here("data", "objects", "zones_filtered.rds"))
```

```{r eval=do_eval}
#| label: process-pols

pols <- sf::st_coordinates(zones_filtered) |> 
  as.data.frame()

saveRDS(pols, here::here("data", "objects", "zones_pols.rds"))
```

```{r eval=do_eval}
#| label: process-map

map <- ggmap::get_stadiamap(bbox = c(left = -90, bottom = 50, right = -83, top = 55.5), 
                            zoom = 5, maptype = "stamen_terrain")

base <- ggmap::ggmap(map) +
  ggplot2::labs(x = "Longitude", y = "Latitude") +
  ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2)

saveRDS(base, here::here("data", "objects", "base_map.rds"))

ont_map <- ggmap::get_stadiamap(bbox = c(left = as.numeric(sf::st_bbox(zones)[1])-1, 
                                         bottom = as.numeric(sf::st_bbox(zones)[2])+1, 
                                         right = as.numeric(sf::st_bbox(zones)[3])-1,
                                         top = as.numeric(sf::st_bbox(zones)[4])+1), 
                            zoom = 5, maptype = "stamen_terrain")

ont_base <- ggmap::ggmap(ont_map) +
  ggplot2::labs(x = "Longitude", y = "Latitude") +
  ggplot2::geom_path(data = pols, ggplot2::aes(x = X, y = Y), col = 1, linewidth = 0.2) +
  ggplot2::geom_rect(
    xmin = -90,
    ymin = 50,
    xmax = -83,
    ymax = 55.5,
    fill = NA, 
    colour = "black",
    size = 0.6
  )

saveRDS(ont_base, here::here("data", "objects", "ont_base_map.rds"))
```

#### Process animal annotations

```{r eval=do_eval}
#| label: process-animals

# Only annotations for animals
# TODO investigate id of cut length 1

anns_wide_animals <- anns_wide_noempty |>
  dplyr::filter(species != 'Human') |> 
  dplyr::filter(species != 'Vehicle') |> 
  dplyr::mutate(id_cut_length = purrr::map_dbl(id , \(x) {
    x |> stringr::str_split("_") |> purrr::map_dbl(~length(.x))
  })) |> 
  dplyr::filter(id_cut_length > 1) |> 
  dplyr::mutate(camera_id = purrr::map_chr(id , \(x) {
    x |>
      stringr::str_split("_") |>
      purrr::map(~.x[[2]]) |>
      unlist()
  }))

# Join to EXIF
# TODO streamline exif so you dont have to join

anns_wide_animals_joined <- anns_wide_animals |>
  dplyr::left_join(exif, by = c("source_file", "image_id"))

stopifnot(dim(anns_wide_animals)[1] ==  dim(anns_wide_animals_joined)[1])
stopifnot((anns_wide_animals_joined$source_file |> is.na() |> sum()) == 0)

anns_wide_animals_joined_mut <- anns_wide_animals_joined |> 
  dplyr::filter(species != "Unknown") |>
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) |> 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_"))

anns_wide_animals_joined_mut[
  anns_wide_animals_joined_mut$ymd=="2024_Aug_19",]$date_time <- lubridate::as_date("2022-08-19")
anns_wide_animals_joined_mut[
  anns_wide_animals_joined_mut$ymd=="2024_Aug_19",]$ymd <- "2022_Aug_19"

saveRDS(anns_wide_animals_joined_mut, here::here("data", "objects", "annotations_animals.rds"))

## Spare code

# Unknowns and bird breakdown
# anns_wide_animals |> dplyr::filter(species == "Unknown") |> 
#   dplyr::pull(id) |> 
#   stringr::str_sub(1,4) |> table() |> sort() 
# anns_wide_animals |> dplyr::filter(stringr::str_detect(species,"Bird")) |> 
#   dplyr::pull(id) |> 
#   stringr::str_sub(1,4) |> table() |> sort()

# anns_wide_animals_joined_dets <- anns_wide_animals |> 
#   dplyr::right_join(dets_images, by = "image_id")

# anns_wide_animals_joined_mut$species |> unique() |> grepl(pattern='sp', fixed = T) |> sum()
```

#### Process deployments from exif data

```{r eval=do_eval}
#| label: process-deployments

camera_days <- exif |> 
  dplyr::group_by(maker_notes_user_label) |> 
  dplyr::summarise(camera_first_time = min(exif_date_time_original)) |> 
  dplyr::mutate(date_time = lubridate::ymd_hms(camera_first_time, 
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_")) |> 
  dplyr::select(camera_id = maker_notes_user_label, 
                camera_first_day = date_time)

camera_days[camera_days$camera_first_day == "2024-05-23 13:31:46",]$camera_first_day <- as.Date("2022-05-23 13:31:46")

camera_regions <- cams_sf_coords |> 
  sf::st_join(zones_filtered) |> 
  sf::st_drop_geometry() |> 
  dplyr::select(camera_id, ECOZONE_NA)

saveRDS(camera_days,  here::here("data", "objects" ,"camera_days.rds"))
saveRDS(camera_regions,  here::here("data", "objects" ,"camera_regions.rds"))
```

#### Process presence absence data

First, absolute occurrence.

```{r eval=do_eval}
#| label: process-pa

anns_wide_animals_table <- anns_wide_animals_joined_mut |>
  dplyr::select(camera_id, species) |> 
  dplyr::filter(species != "Unknown") |>
  table() 
anns_wide_animals_table <- anns_wide_animals_table > 0
anns_wide_animals_table <- anns_wide_animals_table + 0
anns_wide_animals_table <- anns_wide_animals_table |> 
  as.data.frame() |> 
  tibble::rownames_to_column("camera_id") |> 
  dplyr::arrange(camera_id)
anns_wide_animals_table_tb <- anns_wide_animals_table |> 
  dplyr::select(-camera_id)
rownames(anns_wide_animals_table_tb) <- anns_wide_animals_table$camera_id

saveRDS(anns_wide_animals_table_tb,  here::here("data", "objects" ,"anns_wide_animals_table_tb.rds"))
```

#### Process land cover data

1 km buffers

```{r eval=do_eval}
#| label: process-land-cover
lu_16 <- raster::raster(here::here("data-raw", "spatial", "land_use", "Class", "FarNorth_LandCover_Class_UTM16.tif"))
lu_17 <- raster::raster(here::here("data-raw", "spatial", "land_use", "Class", "FarNorth_LandCover_Class_UTM17.tif"))

lu_dat <- readr::read_csv(here::here("data-raw", "spatial", "land_use", "attr_table_northen_ont_lc.txt")) |> 
  dplyr::mutate(cats = as.factor(code))

cams_sf_buffers_16 <- cams_sf_coords |> 
  sf::st_transform(sf::st_crs(lu_16)) |> 
  sf::st_buffer(dist = 1000)

cams_sf_buffers_17 <- cams_sf_coords |> 
  sf::st_transform(sf::st_crs(lu_17)) |> 
  sf::st_buffer(dist = 1000)

extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
                                        progress = FALSE)
extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
                                        progress = FALSE)

ind_16 <- which(lapply(extr_16, function(x) all(is.na(x[,1]))) |> unlist())
ind_17 <- which(lapply(extr_17, function(x) all(is.na(x[,1]))) |> unlist())

stopifnot(length(unique(c(ind_16, ind_17))) == length(extr_16))

extr_16[ind_16] <- extr_17[ind_16]

extr <- extr_16

all_table <- dplyr::bind_rows(extr) |> 
  dplyr::group_by(value) |> 
  dplyr::summarise(total = sum(coverage_fraction)) |> 
  dplyr::mutate(value = as.factor(value), 
                Percentage = round(total/sum(total)*100, 2))

extr_list <- extr |> 
  lapply(FUN = function(x){
    the_tab <- table(x$value)
    props <- the_tab/sum(the_tab)
    cats <- names(the_tab)
    df <- tibble::tibble(cats, props)
    return(df)
  })
names(extr_list) <- cams_sf_buffers_16$camera_id

extr_table <- dplyr::bind_rows(extr_list, .id = "camera_id") |> 
  dplyr::mutate(props = as.double(props)) |> 
  dplyr::left_join(lu_dat, by = "cats") |> 
  dplyr::filter(!is.na(category_code)) |> 
  dplyr::select(camera_id,category_code, props) |> 
  tidyr::pivot_wider(names_from = "category_code", values_from = "props") |> 
  as.data.frame()
extr_table[is.na(extr_table)] <- 0

# Combine the categories
extr_table_simple <- extr_table |> 
  dplyr::mutate(WAT = WAT + XWAT,
                MixTRE = MixTRE + DecTRE,
                Dist  = TrOrSHr + NSWood, 
                SWA = ThSWA + ConSWA, 
                Anth = URB + MIN + BED) |> 
  dplyr::select(-XWAT,
                -DecTRE,
                -TrOrSHr, -NSWood, 
                -ThSWA, -ConSWA, 
                -URB, -MIN, -BED)

extr_table_sub <- extr_table_simple |> 
  dplyr::filter(camera_id %in% anns_wide_animals_table$camera_id) |> 
  dplyr::arrange(camera_id)
extr_table_sub_tb <- extr_table_sub |> 
  dplyr::select(-camera_id)
rownames(extr_table_sub_tb) <- extr_table_sub$camera_id
extr_table_sub_tb <- extr_table_sub_tb[, colSums(extr_table_sub_tb) != 0]

stopifnot(all(rownames(anns_wide_animals_table_tb) == 
                rownames(extr_table_sub_tb)))

saveRDS(extr_table_sub_tb,  here::here("data", "objects" ,"extr_table_sub_tb.rds"))

table_cat <- lu_dat |> 
  dplyr::left_join(all_table, by = dplyr::join_by(cats == value)) |> 
  dplyr::mutate(Percentage = ifelse(is.na(Percentage), 0, Percentage)) |>  
  dplyr::select(category_code, label, Percentage) |> 
  dplyr::filter(category_code != "ND") |> 
  dplyr::arrange(category_code) |> 
  gt::gt() |> 
  gt::cols_label(
    label = "Land cover class",
    category_code = "Abbreviations",
    Percentage = "% Cover in 1km buffers"
  ) |> 
  gt::opt_stylize(style = 1, color = "gray")

saveRDS(table_cat,  here::here("data", "objects" ,"table_cat.rds"))
```

Then, based on events, assuming day as an event window.

```{r eval=do_eval}
#| label: events-Plots

events_days <- anns_wide_animals_joined_mut |> 
  dplyr::select(plot_id, species, month, ymd) |> 
  dplyr::distinct()

events_summarised <- events_days |> 
  dplyr::group_by(plot_id, species) |>
  dplyr::summarise(day_count = dplyr::n()) |> 
  dplyr::ungroup()

events_tb <- events_summarised |> 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
  dplyr::arrange(plot_id) |>
  tibble::column_to_rownames("plot_id")

events_tb[is.na(events_tb)] <- 0
events_tb_filt <- events_tb[,colSums(events_tb) > 50]
tmp_col <- ncol(events_tb_filt)

events_tb_filt_with_col <- events_tb_filt |> 
  tibble::rownames_to_column("plot_id")

saveRDS(events_tb_filt_with_col,  here::here("data", "objects" ,"events_tb_filt_with_col.rds"))
```

Same, but using camera instead of plot id.

```{r eval=do_eval}
#| label: events-Cams

events_cam_days <- anns_wide_animals_joined_mut |> 
  dplyr::select(camera_id, species, month, ymd) |> 
  dplyr::distinct()

saveRDS(events_cam_days,  here::here("data", "objects" ,"events_cam_days.rds"))

events_cam <- events_cam_days |> 
  dplyr::group_by(camera_id, species) |>
  dplyr::summarise(day_count = dplyr::n()) |> 
  dplyr::ungroup()

events_tb_cam <- events_cam |> 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") |>
  dplyr::arrange(camera_id) |>
  tibble::column_to_rownames("camera_id")

events_tb_cam[is.na(events_tb_cam)] <- 0
# events_tb_cam_filt <- events_tb_cam[,colSums(events_tb_cam)>50]
# tmp_col <- ncol(events_tb_cam)

saveRDS(events_tb_cam,  here::here("data", "objects" ,"events_tb_cam.rds"))
```

#### Process scientific names

```{r eval=do_eval}
# sci_names <- taxize::comm2sci(sp_tab$species)
sci_names <- readRDS(here::here("data", "objects", "sci_names.rds")) |> 
  unlist() |> tibble::as_tibble(rownames = NA) |> 
  tibble::rownames_to_column() |> 
  dplyr::rename(Scientific_name = value,
                species = rowname) |> 
  tibble::add_row(Scientific_name = "Ursus americanus", species = "Black bear") |> 
  tibble::add_row(Scientific_name = "Perisoreus canadensis", species = "Canada Jay")

saveRDS(sci_names,here::here("data", "objects" , "sci_names_processed.rds"))
```

#### Process WildTrax data

```{r}
# debugonce(wt_auth)
wildRtrax::wt_auth(force = T)

# debugonce(wt_get_download_summary)
my_projects <- wildRtrax::wt_get_download_summary(sensor_id = 'CAM')

# debugonce(wt_download_report)
my_report <- wildRtrax::wt_download_report(project_id = 2431, sensor_id = 'CAM', reports = "image_report", weather_cols = F) |>
  tibble::as_tibble()
```